---
title: "HW 3 for P8105 - hlz2108"
author: "Helen Zhang"
date: "October 5, 2020"
output: github_document
---
Below is the code for setup:

```{r setup}
library(tidyverse)
library(dplyr)
library(readxl)
library(ggplot2)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 0

This “problem” focuses on structure of your submission, especially the use git and GitHub for reproducibility, R Projects to organize your work, R Markdown to write reproducible reports, relative paths to load data from local files, and reasonable naming structures for your files.

## Problem 1

```{r instacart_df}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns.

Observations are the level of items in order by user. 

How many aisles, and which are most items ordered from?

```{r aisle_count}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Let's make a plot

```{r instacart_plot}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle,)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```




How many aisles are there, and which aisles are the most items ordered from?
Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

## Problem 2

```{r accel_df}
accel_df = 
  read_csv("./data/accel_data.csv", col_types = cols(
      week = col_integer(),
      day_id = col_integer()
      )) %>%
  janitor::clean_names() %>% 
  mutate(
    weekend = case_when(
      day == "Monday" ~ "weekday",
      day == "Tuesday" ~ "weekday",
      day == "Wednesday" ~ "weekday",
      day == "Thursday" ~ "weekday",
      day == "Friday" ~ "weekday",
      day == "Saturday" ~ "weekend",
      day == "Sunday" ~ "weekend"
)) %>% 
  relocate("weekend") %>%
  pivot_longer(
    activity_1:activity_1440, 
    names_to = "activity",
    values_to = "activity_count") %>% 
  separate(activity, into = c("activity", "activity_minute")) %>%
  select(-activity) %>%
  mutate(activity_minute = as.integer(activity_minute)) %>% view
```

The accel data set contains `r nrow(accel_df)` rows and `r ncol(accel_df)` columns. The dataset contains 5 weeks of accelerometer data collected from a 63 year-old male with BMI 25, who was admitted to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed with congestive heart failure (CHF). The variables in this dataset are `r names(accel_df)`. Below is a detailed description of several variables of interest:

* _day_id_ is a `r class(pull(accel_df,day_id))` variable and indicates the day the accelerometer data was collected.

* _week_ is a `r class(pull(accel_df, week))` variable and indicates the week the accelerometer data was collected.

* _weekend_ is a `r class(pull(accel_df, weekend))` variable that indicates whether the information was collected on a weekend vs. a weekday.

* _activity_minute_ is a `r class(pull(accel_df, activity_minute))` variable that indicates the minute when the activity count was collected, corresponding to each minute of a 24-hour day starting at midnight.

```{r aggregate_accel}
aggregate_accel_df = accel_df %>% 
  group_by(day_id, day) %>%
  summarize(total_activity = sum(activity_count)) %>% 
  knitr::kable() %>% view
```

Activity counts were at their lowest on the two Saturdays prior to end of data collection (weeks 4 and 5).

```{r accel_plot}
accel_df %>% 
  ggplot(aes(x = activity_minute, y = activity_count, color = day)) + 
  geom_line() +
  labs(
    title = "24-Hour Activity Count by Day",
    x = "Time",
    y = "Activity Count",
    caption = "Data from the accel dataset"
  ) + 
  scale_x_continuous(
    breaks = c(0, 360, 720, 1080, 1440), 
    labels = c("12AM", "6AM", "12PM", "6PM", "11:59PM"),
    limits = c(0, 1440)
    ) + 
    viridis::scale_color_viridis(
      name = "Day",
      discrete = TRUE) + 
  theme(legend.position = "bottom")
```

## Problem 3

```{r ny_noaa_df}
library(p8105.datasets)
data("ny_noaa")
```

